\section{Entry Point (main.cpp)}

\subsection{Purpose}

\texttt{src/main.cpp} is a minimal entry point.
It contains no business logic: it delegates argument parsing to the \texttt{args} module
and all compression/decompression work to the \texttt{codec} module.
Keeping it thin means the codec is independently testable --- it only requires
\texttt{std::istream} and \texttt{std::ostream} references, not real files.

\subsection{Implementation}

The function performs exactly four logical steps:

\begin{enumerate}
  \item \textbf{Parse arguments:}
        \texttt{ParsedArgs args = parse\_arguments(argc, argv)}.
        On invalid arguments, \texttt{parse\_arguments} prints a usage message and exits;
        \texttt{main} never sees an invalid \texttt{ParsedArgs}.

  \item \textbf{Open streams:}
        \texttt{std::ifstream in(args.infile, std::ios::binary)} and
        \texttt{std::ofstream out(args.outfile, std::ios::binary)}.
        Both are checked immediately; if either fails, an error is printed to
        \texttt{std::cerr} and \texttt{main} returns 1.

  \item \textbf{Dispatch:}
        \texttt{args.decompress}
        $\to$ \texttt{decompress(in, out)}
        or \texttt{compress(in, out, args.width, args.use\_model, args.adaptive\_scan)}.

  \item \textbf{Return:} the exit code from the called function (0 = success, 1 = error).
\end{enumerate}

\subsection{DEBUG Macro}

When compiled with \texttt{-DDEBUG}, \texttt{main.cpp} prints mode, input/output paths,
image width, and the model/scan flags to \texttt{std::cout} before dispatching.
This output is conditionally compiled:

\begin{lstlisting}[style=cpp]
#ifdef DEBUG
    std::cout << "Mode: " << (args.decompress ? "decompress" : "compress") << "\n";
    // ...
#endif
\end{lstlisting}

No debug output is produced in release builds.
The \texttt{Makefile} does not define \texttt{DEBUG} by default.

\subsection{Error Messages: Czech vs English}

Several error strings in \texttt{main.cpp} and \texttt{codec.cpp} are currently in Czech
(e.g., \textit{``Nelze otevřít vstupní soubor:''}, \textit{``Neplatný formát souboru''}).
The assignment requires English \emph{comments} in source files; error messages to
\texttt{stderr} are not comments and are therefore technically not in scope of that rule.
However, mixing Czech runtime output with an otherwise English-commented codebase is
inconsistent.
See the Improvements chapter for the recommended fix.
